---
title: "PredictionAssignment"
author: "mansipriya"
date: "September 18, 2018"
output: html_document
---
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to
collect a large amount of data about personal activity relatively inexpensively. These
type of devices are part of the quantified self movement - a group of enthusiasts who
take measurements about themselves regularly to improve their health, to find patterns 
in their behavior, or because they are tech geeks. One thing that people regularly do is
quantify how much of a particular activity they do, but they rarely quantify how well 
they do it. In this project, my goal will be to use data from accelerometers on the 
belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell 
lifts correctly and incorrectly in 5 different ways. More information is available 
from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the
Weight Lifting Exercise Dataset)

### Getting the data
```{r}
training_url <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
test_url <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
training <- read.csv(url(training_url))
test <- read.csv(url(test_url))
```
### Analyzing the data

```{r}
dim(training)
head(training)
dim(test)
head(test)
```

We can see lot of NAs are present in the dataset. Lets first remove that

```{r}
training <- training[, colSums(is.na(training)) == 0]
test <- test[, colSums(is.na(test)) == 0] 
## Removing unnecessary data
classe<- training$classe
trainRemove<- grepl("^X|timestamp|window", names(training))
training<- training[, !trainRemove]
training<- training[, sapply(training, is.numeric)]
training$classe <- classe
testRemove<- grepl("^X|timestamp|window", names(test))
test<- test[, !testRemove]
test<- test[, sapply(test, is.numeric)]
```
### Explopratory Data Analysis
```{r}
class_table <- data.frame(table(training$classe))
library(ggplot2)
qplot(x = class_table$Var1, y = class_table$Freq,data = class_table, geom = "point", xlab = "Activity", ylab = "Count")
```
We can see Activity A is the most Frequent activity done by people

Now lets create training and test set
```{r}
library(caret)
set.seed(12345)
inTrain <- createDataPartition(training$classe, p=0.70, list=F)
train_data <- training[inTrain, ]
test_data <- training[-inTrain, ]
```

### Model Training
```{r}
## Lets firt try decision tree
fit1 <- train(classe~.,data = train_data, method = "rpart")
library(rattle)
fancyRpartPlot(fit1$finalModel)
```

Lets check accuracy of this model
```{r}
pred <- predict(fit1,test_data)
confusionMatrix(test_data$classe , pred)
```

As we can see accuracy is quite low. lets try with another model

```{r}
fit2 <- train(classe ~., data=train_data, method="rf",  trControl=trainControl(method="cv", 5), ntree=100)
pred <- predict(fit2,test_data)
confusionMatrix(test_data$classe , pred)
```

In this model we get  accuracy around 99%. Hence this model is a good one to rely on for predictions.

## Predicting for test data

```{r}
test_pred <- predict(fit2 , test)
test_pred
```